---
title: "253 Final Project"
author: "Melissa Blum, Andrew Padgett, Tomas Panek"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r, echo=FALSE}
#plotting and exploring
library(tidyverse) #for plotting and summarizing
library(GGally) #for nice scatterplot matrix 
library(ggridges) #for joy/ridge plots
library(corrplot) #for basic correlation matrix plot
library(naniar) #for exploring missing values
library(pdp) #for partial dependence plots, MARS models
library(rpart.plot) #for plotting decision trees
library(vip) #for importance plots
library(pROC) #for ROC curves
library(plotROC) #for plotting ROC curves

#making things look nice
library(lubridate) #for nice dates
library(knitr) #for nice tables
library(scales) #for nice labels on graphs
library(gridExtra) #for arranging plots
library(broom) #for nice model output
library(janitor) #for nice names

#data
library(ISLR) #for data
library(moderndive) #for data
library(rattle) #weather data

#modeling
library(rsample) #for splitting data
library(recipes) #for keeping track of transformations
library(caret) #for modeling
library(leaps) #for variable selection
library(glmnet) #for LASSO
library(earth) #for MARS models
library(rpart) #for decision trees
library(randomForest) #for bagging and random forests

theme_set(theme_minimal())
```

I read in the diabetes hospital readmission data in case that's what we end up using. There's one paper that's been published using the same data, so I put the link to that paper below. They used a logistic regression model with many of the variables in the dataset to predict hospital readmissions of diabetic patients - I think it would be cool to try other types of models and see if we can find one that makes more accurate predictions.

archive.ics.uci.edu/ml/machine-learning-databases/00296/

```{r}
diabetic_data <- read_csv("diabetic_data.csv", 
    na = "?")
```

```{r}
head(diabetic_data)
```

Links to data and paper:

http://archive.ics.uci.edu/ml/datasets/Diabetes+130-US+hospitals+for+years+1999-2008#
https://www.hindawi.com/journals/bmri/2014/781670/
